---
title: "Problem4"
output: html_document
---
library(ggplot2)
data(diamonds)
names(diamonds)
### price vs. x
```{r}
ggplot(aes(x=price, y=x), data=diamonds) + geom_point() + geom_smooth(method='lm', color='red')
```
***

### What are your observation about the scatterplot of price vs x?
Notes:
Price seems to be related to x
***

### Correlations price and x
```{r}
with(diamonds, cor.test(price, x))
```

***

### Correlations price and y
```{r}
with(diamonds, cor.test(price, y))
```

***

### Correlations price and z
```{r}
with(diamonds, cor.test(price, z))
```

***

### price vs depth
```{r}
ggplot(aes(x=price, y=depth), data=diamonds) + geom_point()
```

### Adjustments - price vs. depth
range(diamonds$depth)
?scale_x_continuous
```{r}
ggplot(aes(x=depth, y=price), data=diamonds) + geom_point(alpha=1/100) + scale_x_continuous(breaks=seq(43,79,2))
```

***

### Correlations depth and price

```{r}
with(diamonds, cor.test(depth, price))
```

***

### price vs carat
```{r}
ggplot(aes(x=carat, y=price), data=diamonds) + geom_point() + 
  xlim(0, quantile(diamonds$carat, 0.99)) + 
  ylim(0, quantile(diamonds$price, 0.99))
```

***

### price vs. volume
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z 
head(diamonds, 20)
ggplot(aes(x=diamonds$volume, y=price), data=diamonds) + geom_point()
```
***

### Findings - price vs. volume

```{r}
library('dplyr')
library('plyr')
detach('package:plyr', unload=TRUE)
count(diamonds$volume == 0)
```

***

### Correlations on Subsets

```{r}
with(subset(diamonds, volume>0&volume<800), cor.test(price, volume))
clean_diamonds <- subset(diamonds, volume>0&volume<800)
```

***

### Adjustments - price vs. volume
```{r}
ggplot(aes(x=clean_diamonds$volume, y=price), data=clean_diamonds) + geom_point(alpha=1/10) + geom_smooth(method='lm', color='red')
```

***

### Mean Price by Clarity
```{r}
detach("package:plyr", unload=TRUE)
diamonds.diamondsByClarity <- diamonds %>% group_by(clarity) %>% summarize( 
                          mean_price = mean(as.numeric(price)),
                          median_price = median(as.numeric(price)),
                          min_price = min(as.numeric(price)),
                          max_price = max(as.numeric(price)),
                          n = n()) %>% arrange(clarity)
head(diamonds.diamondsByClarity, 20)
```

***

### Bar Charts of Mean Price

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```

***