White Wine Quality Exploration by Josh See
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

### Load the Data

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:\\Users\\Josh\\Documents\\GitHub\\P3-Data-Analysis-with-R")
#setwd('/Volumes/MacHD/Projects/P3-Data-Analysis-with-R')
wines <- read.csv('wineQualityWhites.csv')
# Selecting only col 2 to col 13, excluding the first col (Line no.)
wines <- wines[,c(2:13)]
```

```{r echo=FALSE}
dim(wines)
summary(wines)
names(wines)
str(wines)
```

# Univariate Plots Section

## Observation and summary
We already know that the quality of the wine is rated by wine experts from the scale of 0 (very bad) to 10 (very excellent)

The lowest and highest quality score given is 3 and 9, with a mean of 5.878

All the attribute provided is decimal values except quality which is a integer.

## Lets explore each individual variables first

### quality
```{r echo=FALSE}
ggplot(aes(x = quality), data = wines) + geom_histogram(binwidth = 1)   
```

```{r echo=FALSE}
# Since wine is just an integer, change that to a categorical variables
wines$quality <- factor(wines$quality)
summary(wines$quality)
```

### fixed.acidity
```{r echo=FALSE}
ggplot(aes(x = fixed.acidity), data = wines) + geom_histogram(binwidth = .2)
```

### volatile.acidity
```{r echo=FALSE}

```

### citric.acid
```{r echo=FALSE}
ggplot(aes(x = citric.acid), data = wines) + geom_histogram(binwidth = 0.05)
```

### residual.sugar
```{r echo=FALSE}
ggplot(aes(x = residual.sugar), data = wines) + geom_histogram(binwidth = 0.2)
```

### chlorides
```{r echo=FALSE}

```

### free.sulfur.dioxide
```{r echo=FALSE}

```

### total.sulfur.dioxide
```{r echo=FALSE}

```

### pH
```{r echo=FALSE}
ggplot(aes(x = pH), data = wines) + geom_histogram(binwidth = 0.02)
```

### sulphates
```{r echo=FALSE}
ggplot(aes(x = sulphates), data = wines) + geom_histogram(binwidth = 0.02)
```

### density
```{r echo=FALSE}
table(wines$density)
ggplot(aes(x = density), data = wines) + geom_histogram(binwidth = 0.0005) + coord_cartesian(xlim=c(0.98,1.01))+ scale_x_continuous(breaks=seq(0.98,1.01,.005))

summary(wines$density)
```

### alcohol
```{r echo=FALSE}
ggplot(aes(x=alcohol), data = wines) +
  geom_bar(binwidth = 0.2) +
  xlab("Alcohol (% by volume)") +
  ylab("Occurence") +
  scale_x_continuous(limits = c(8,15), breaks = seq(8,15,1)) +
  ggtitle("Histogram for Alcohol \n")
```


# Univariate Analysis

### What is the structure of your dataset?

There are 4,898 observations and 13 variables in the dataset. First variable is basically an index of each observation. I removed it as its useless for the analysis. The last variable which is an integer quality score and the rest of the variables are stored as numerical data types. The quality score is rated by wine experts from the scale of 0 (very bad) to 10 (very excellent). The quality score is converted to a factor data types. Most of the wines have a quality score of 6, the lowest and highest score given is 3 and and 9.

### What is/are the main feature(s) of interest in your dataset?



### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

ggplot(aes(x=factor(quality)), data = wines) + geom_histogram()

ggplot(aes(x=fixed.acidity), data=wines) + geom_histogram(binwidth=0.2)
ggplot(aes(x=volatile.acidity), data=wines) + geom_histogram(binwidth=0.02)
ggplot(aes(x=citric.acid), data=wines) + geom_histogram(binwidth=0.02)
ggplot(aes(x=residual.sugar), data=wines) + geom_histogram(binwidth=0.2)
ggplot(aes(x=chlorides), data=wines) + geom_histogram(binwidth=0.01)

ggplot(aes(x=alcohol), data=wines) + geom_bar(binwidth=0.1) + scale_x_continuous(limits = c(8,15), breaks = seq(8,15,1))
summary(wines$alcohol)

summary(wines$citric.acid)
