White Wine Quality Exploration by Josh See
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
```

### Load the Data

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:\\Users\\Josh\\Documents\\GitHub\\P3-Data-Analysis-with-R")
wines <- read.csv('wineQualityWhites.csv')
# Selecting only col 2 to col 13, excluding the first col (Line no.)
modifiedwines <- wines[,c(2:13)]
```

```{r echo=FALSE}
dim(wines)
summary(wines)
names(wines)
str(wines)
```

## Summary

The lowest and highest quality score given is 3 and 9, with a mean of 5.878

All the attribute provided is decimal values except quality which is a integer.


ggplot(aes(x=factor(quality)), data = wines) + geom_histogram()

ggplot(aes(x=fixed.acidity), data=wines) + geom_histogram(binwidth=0.2)
ggplot(aes(x=volatile.acidity), data=wines) + geom_histogram(binwidth=0.02)
ggplot(aes(x=citric.acid), data=wines) + geom_histogram(binwidth=0.02)
ggplot(aes(x=residual.sugar), data=wines) + geom_histogram(binwidth=0.2)
ggplot(aes(x=chlorides), data=wines) + geom_histogram(binwidth=0.01)

ggplot(aes(x=alcohol), data=wines) + geom_bar(binwidth=0.1) + scale_x_continuous(limits = c(8,15), breaks = seq(8,15,1))
summary(wines$alcohol)

summary(wines$citric.acid)
