White Wine Quality Exploration by Josh See
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
library(scales)
```
 
### Introduction
The goal of this project is to explore the chemical properties found in a tidy data set of white wines and to understand and summarize which properties is closely related to the wine quality score.

### Load the Data
```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
setwd("C:\\Users\\Josh\\Documents\\GitHub\\P3-Data-Analysis-with-R")
#setwd('/Volumes/MacHD/Projects/P3-Data-Analysis-with-R')
wines <- read.csv('wineQualityWhites.csv')
```

### Data summary 
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_summary}
# Data Summary
summary(wines)
```

### Data dimensions
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_dimensions}
# Data Dimensions
dim(wines)
```

### Feature names
```{r echo=FALSE, message=FALSE, warning=FALSE, Feature_names}
# Feature Names
names(wines)
```

### Data Structure
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_structure}
str(wines)
```

I found out that the first col X is a unique index variable for each individual observation. I think it is not very useful to the analysis so its best to remove it before procceding to the next step.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Selecting only col 2 to col 13, excluding the first col (Line no.)
wines <- wines[,c(2:13)]
```

# Univariate Plots Section

### Observation
We already know that the quality of the wine is rated by wine experts from the scale of 0 (very bad) to 10 (very excellent)

The lowest and highest quality score given is 3 and 9, with a mean of 5.878

All the attribute provided is decimal values except quality which is a integer.

Attributes like fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, sulphates have max value that is greater than 75% quantile.

Lets take a look at each properties to get a sense of data distribution.

### Quality
```{r echo=FALSE, message=FALSE, warning=FALSE, Quality}
# Since wine quality score is just an integer, we can change quality scores as a factor, so that we can use it later
wines$quality_as_factor <- factor(wines$quality, levels=c(0,1,2,3,4,5,6,7,8,9,10))
```

#### Levels attributes of wine quality
```{r echo=FALSE, message=FALSE, warning=FALSE, Levels_attributes_of_wine_quality}
levels(wines$quality_as_factor)
```

#### Summary
```{r echo=FALSE, message=FALSE, warning=FALSE, summary}
summary(wines$quality)
```
#### Summary quality as factor
```{r echo=FALSE, message=FALSE, warning=FALSE, summary}
summary(wines$quality_as_factor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, summary}
qplot(x=quality_as_factor, data = wines, binwidth=1, 
      xlab='Quality', 
      ylab='Count',
      color=I('black'), fill=I('#F79420'))
```

I factor the wine quality from an integer variable to a categorical varible as its kind of arbitary and can be represented by another form of values not just integer. The histogram shows the shape of normal distribution. 

The range of possible scores is from 0 to 10, in the dataset the minimum score is 3 and maximum is 9. The mean is 5.878 and the median is 6 which is very close to each other.

```{r echo=FALSE, message=FALSE, warning=FALSE}
Wine_quality_mode <- names(which.max(table(wines$quality)))
```
So the mode of the quality score is 6.

### Fixed.acidity
```{r echo=FALSE, message=FALSE, warning=FALSE, fixed_acidity}
table(wines$fixed.acidity)
summary(wines$fixed.acidity)
# Box plot to visualize the outlier
qplot(y=fixed.acidity,  x = factor(0), data=wines, geom='boxplot')

qplot(x=fixed.acidity, data=wines, binwidth=.2, 
      xlab='Fixed Acidity', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(lim=c(4, 11))
```

There seems be a right tail when the histogram was initially chartted. Once the outlier which is a value 11.8 and 14.2 is removed, the histogram show a normal distribution.

### Volatile.acidity
```{r echo=FALSE, message=FALSE, warning=FALSE, volative_acidity}
summary(wines$volatile.acidity)

summary(log10(wines$volatile.acidity))

qplot(x=log10(volatile.acidity), data=wines, binwidth=.03,
      xlab='Volatile Acidity', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) 
```

By transforming the variable using a log based 10, we able to remove the right tail and turn it into a normal distribution.

### Citric.acid
```{r echo=FALSE, message=FALSE, warning=FALSE, citric_acid}
summary(wines$citric.acid)
# Box plot to visualize the outlier
qplot(y=citric.acid,  x=factor(0), data=wines, geom='boxplot')

qplot(x=citric.acid, data=wines,
      xlab='Citric Acidity', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(lim=c(0, 1))
```

Using a box plot, we can cleary visualize the outlier which may cause the histogram to have a long right tail. Once we removed that, our histogram shows a normal distribution.

### Residual.sugar
```{r echo=FALSE, message=FALSE, warning=FALSE, residual_sugar}
summary(wines$residual.sugar)

summary(log10(wines$residual.sugar))

qplot(y=residual.sugar,  x=factor(0), data=wines, geom='boxplot')

# Removing outlier on the right tail
qplot(x=residual.sugar, data = wines,
      xlab='Residual Sugar', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(lim=c(0, 40))

# Transform the variable
qplot(x=log10(residual.sugar), data = wines,
      xlab='Residual Sugar', 
      ylab='Count',
      color=I('black'), fill=I('#F79420'))
```

I try to remove the outlier after looking at the boxplot. The histogram seems to show a positive skewed graph. However, after applying a log based 10 transformation. This histogram still does not appear to be a normal distribution. It looks more like a bi-model distribution.

### chlorides
```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides}
summary(log10(wines$chlorides))

qplot(x=log10(wines$chlorides), data=wines,
      xlab='Chlorides', 
      ylab='Count',
      color=I('black'), fill=I('#F79420'))

```

Chlorides shows a normal distribution once a log based 10 transformation is applied.

### free.sulfur.dioxide
```{r echo=FALSE, message=FALSE, warning=FALSE, free_sulfur_dioxide}
summary(wines$free.sulfur.dioxide)

qplot(y=free.sulfur.dioxide,  x=factor(0), data=wines, geom='boxplot')

qplot(x=free.sulfur.dioxide, data=wines, binwidth=2,
      xlab='Free Sulfur Dioxide', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(lim=c(0, 90))
```

Once outlier is removed, the histogram appears as a normal distribution.

### total.sulfur.dioxide
```{r echo=FALSE, message=FALSE, warning=FALSE, total_sulfur_dioxide}
summary(wines$total.sulfur.dioxide)

qplot(y=total.sulfur.dioxide,  x=factor(0), data=wines, geom='boxplot')

qplot(x = total.sulfur.dioxide, data = wines, , binwidth=5, 
      xlab='Total Sulfur Dioxide', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(lim = c(0, 300))
```

Same for total.sulfur.dioxide once right tail outlier is removed, the histogram appears as a normal distribution.

### pH
```{r echo=FALSE, message=FALSE, warning=FALSE, ph}
summary(wines$pH)

qplot(x = pH, data=wines, binwidth = .05,
      xlab='pH', 
      ylab='Count',
      color=I('black'), fill=I('#F79420'))
```

pH seems to give a normal distribution once histogram is graphed. There is no transformation that needs to be applied to it.

### sulphates
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(wines$sulphates)

qplot(y=sulphates,  x=factor(0), data=wines, geom='boxplot')

summary(log10(wines$sulphates))

qplot(y=log10(sulphates),  x=factor(0), data=wines, geom='boxplot')

qplot(x = log10(sulphates), data=wines, binwidth = .02,
      xlab='Sulphates', 
      ylab='Count',
      color=I('black'), fill=I('#F79420'))
```

Applying log based 10 transformation gives us a normal distribution.

### density
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(wines$density)

qplot(y=density,  x=factor(0), data=wines, geom='boxplot')

qplot(x = density, data=wines, binwidth=.0005,
      xlab='Density', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(limits = c(0.98, 1.01), breaks=seq(0.98, 1.01, .005))
  
```

The desity histogram shows a normal distribution once the outlier is removed.

### alcohol
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(wines$alcohol)

tapply(wines$alcohol, wines$quality, mean)

qplot(y=alcohol,  x=factor(0), data=wines, geom='boxplot')

qplot(x = alcohol, data=wines, binwidth=.2, 
      xlab='Alcohol', 
      ylab='Count',
      color=I('black'), fill=I('#F79420')) +
  scale_x_continuous(limits = c(8, 14.20), breaks=seq(8, 14.20, 1))
```

This histogram shows a slight positive skewed distribution with a peak between 9 and 10.

# Univariate Analysis

### What is the structure of your dataset?

There are 4,898 observations and 13 variables in the dataset. First variable X is basically an index of each observations. It is being removed as its useless for the analysis. The rest of the variables which is stored as numerical data type is basically the properties of white wine. Last variable is an integer quality score which is rated by wine experts from the scale of 0 (very bad) to 10 (very excellent). The quality score is converted to a factor data types. Most of the wines have a quality score of 6, the lowest and highest score given is 3 and and 9.

### What is/are the main feature(s) of interest in your dataset?

The main feature that is important to this analysis is the wine quality. Analysis needs to be done on the wine properties to see whether it has an impact to this outcome.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

Lets take a look at how correlated the variables are:

```{r echo=FALSE, message=FALSE, warning=FALSE}
cor(x=wines[,1:11],y=wines$quality)
```

We can see that the following variables are correlated with quallity:
* density (Highly correlated)
* chlorides (Moderatelly correlated)
* volatile.acidity (Moderatelly correlated)
* total.sulfur.dioxide (Moderatelly correlated)
* fixed.acidity (Moderatelly correlated)
* residual.sugar
* citric.acid
* free.sulfur.dioxide
* sulphates
* pH
* alcohol (Highly correlated)

We will concentrate on the top variables that show strong correlation

### Did you create any new variables from existing variables in the dataset?

Yes, since we know that the mode of quality score is 6. We would consider the average is 6 in a 0 to 10 scale. We can define a cut in the scores. 

wines$quality_cut <- cut(wines$quality, breaks=c(0,5,6,10))

wines$quality_cut <- cut(wines$quality, breaks=c(0,4,7,10))

table(wines$quality_cut)

So we will have three groups of wines after the cut. First group has a quality score from 0 to 5, which we can consider the bad quality group. Then the second group which has a score of 6 average quality group and lastly the 7 to 10 quality score group is the best quality group. We will these groups in our analysis.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I created a few histograms to understand the distribution of the features and box plots to find out the outliers. Yes, there were a few outliers in the features in which I removed to get it to look gaussian or normal distribution. I also applied a log based 10 transformation on the features which had long tails so that the features becomes gaussian. Quality and ph were the features which I did not apply any transformations as they look like normal distribution. Alcohol feature looks like a positive skewed distribution.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE}
#set.seed(3000)
#ggpairs(wines[sample.int(nrow(wines), 1500),])
```

I decided to plot out ggpairs to have a look into the relationships. I found out that the 4 most correclated variables with quality were alcohol, ph, chlorides, density. But lets plot all other variables as well.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
p1 <- ggplot(aes(x = quality_as_factor, y = alcohol), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$alcohol), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("Alcohol") +
      ggtitle('Quality vs Alcohol')

p2 <- ggplot(aes(x = quality_as_factor, y = density), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$density), linetype='longdash', alpha=.5, color='blue') +
      ylim(c(0.98, 1.01)) +
      xlab("Quality") +
      ylab("Density") +
      ggtitle("Quality vs Density")

p3 <- ggplot(aes(x = quality_as_factor, y = log10(chlorides)), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(log10(wines$chlorides)), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("log10(Chlorides)") +
      ggtitle("Quality vs log10(Chlorides)")

p4 <- ggplot(aes(x = quality_as_factor, y = log10(volatile.acidity)), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(log10(wines$volatile.acidity)), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("log10(Volatile.acidity)") +
      ggtitle('Quality vs log10(Volatile.acidity)')

p5 <- ggplot(aes(x = quality_as_factor, y = total.sulfur.dioxide), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$total.sulfur.dioxide), linetype='longdash', alpha=.5, color='blue') +
      ylim(c(0, 300)) + 
      xlab("Quality") +
      ylab("Total.sulfur.dioxide") +
      ggtitle('Quality vs Total.sulfur.dioxide')

p6 <- ggplot(aes(x = quality_as_factor, y = fixed.acidity), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$fixed.acidity), linetype='longdash', alpha=.5, color='blue') +
      ylim(c(4, 11)) +
      xlab("Quality") +
      ylab("Fixed.acidity") +
      ggtitle('Quality vs Fixed.acidity')

p7 <- ggplot(aes(x = quality_as_factor, y = log10(residual.sugar)), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(log10(wines$residual.sugar)), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("log10(Residual.sugar)") +
      ggtitle("Quality vs log10(Residual.sugar)")

p8 <- ggplot(aes(x = quality_as_factor, y = citric.acid), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$citric.acid), linetype='longdash', alpha=.5, color='blue') +
      ylim(c(0, 1)) +
      xlab("Quality") +
      ylab("Citric.acid") +
      ggtitle("Quality vs Citric.acid")

p9 <- ggplot(aes(x = quality_as_factor, y = free.sulfur.dioxide), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$free.sulfur.dioxide), linetype='longdash', alpha=.5, color='blue') +
      ylim(c(0, 90)) +
      xlab("Quality") +
      ylab("Free.sulfur.dioxide") +
      ggtitle("Quality vs Free.sulfur.dioxide")

p10 <- ggplot(aes(x = quality_as_factor, y = log10(sulphates)), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(log10(wines$sulphates)), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("log10(Sulphates)") +
      ggtitle("Quality vs log10(Sulphates)")

p11 <- ggplot(aes(x = quality_as_factor, y = pH), data = wines) +
      geom_boxplot() +
      geom_hline(show_guide=T, yintercept=mean(wines$pH), linetype='longdash', alpha=.5, color='blue') +
      xlab("Quality") +
      ylab("pH") +
      ggtitle('Quality vs pH')

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11, ncol = 4)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

mean(wines$alcohol)

tapply(wines$alcohol, wines$quality, mean)

mean(wines$pH)

tapply(wines$pH, wines$quality, mean)

mean(wines$density)

tapply(wines$density, wines$quality, mean)

mean(wines$chlorides)

tapply(wines$chlorides, wines$quality, mean)

```

From the boxplot above, we were able to understand more about the relationship between `quality` and other variables. In the `alcohol` vs `quality` plot, we can clearly see that as the `quality` increases as `alcohol` varies. It has a mean of 10.514 and the mean increases as the quality score is above 5. Same goes with `pH`. It `pH` mean values starts to increase when is above 5. `density` and `chlorides` on the other hand shows a negative influence to `quality` score. Its means are 0.994 and 0.046 for `density` and `chlorides`, both average means decreases as the quality score is above 5.

We can verify this using a scatter plot.
### Quality vs Alcohol and Quality vs pH

```{r echo=FALSE, message=FALSE, warning=FALSE}
p12 <- ggplot(aes(x = quality, y = alcohol), data = wines) + 
      geom_jitter(alpha=1/5) +
      geom_smooth(method='lm') +
      xlab("Quality") +
      ylab("Alcohol")

p13 <- ggplot(aes(x = quality, y = pH), data = wines) + 
      geom_jitter(alpha=1/5) +
      geom_smooth(method='lm') +
      xlab("Quality") +
      ylab("pH")

grid.arrange(p12,p13, ncol = 1)
```

### Quality vs Density and Quality vs Chlorides

```{r echo=FALSE, message=FALSE, warning=FALSE}
p14 <- ggplot(aes(x = quality, y = density), data = wines) + 
      geom_jitter(alpha=1/5) +
      ylim(c(0.98, 1.01)) +
      geom_smooth(method='lm') +
      xlab("Quality") +
      ylab("Density")

p15 <- ggplot(aes(x = quality, y = log10(chlorides)), data = wines) + 
      geom_jitter(alpha=1/5) +
      geom_smooth(method='lm') +
      xlab("Quality") +
      ylab("Chlorides")

grid.arrange(p14,p15, ncol = 1)
```

We already see a tendency in the boxplot, this can be better illustrate with a scatter plot and a linear regression line in it. Good wines tends to have higher alcohol level and higher pH.  

We explore further to find out the influence of main features with other secondary features.

### Alcohol vs Other Features

```{r echo=FALSE, message=FALSE, warning=FALSE}

p16 <- ggplot(aes(x = alcohol, y = pH), data = wines) + 
       geom_jitter(alpha=1/5) +
       geom_smooth(method='lm') + 
       ggtitle("pH vs Alcohol") +
       xlab("Alcohol") +
       ylab("pH")

p17 <- ggplot(aes(x = alcohol, y = density), data = wines) + 
       geom_jitter(alpha=1/5) +
       geom_smooth(method='lm') +
       ylim(c(0.985, 1.005)) +  
       ggtitle("Density vs Alcohol") +
       xlab("Alcohol") +
       ylab("Density")

p18 <- ggplot(aes(x = alcohol, y = log10(residual.sugar)), data = wines) + 
       geom_point(alpha = 1/5) +         
       geom_smooth(method = 'lm') + 
       ggtitle('Log10(Residual Sugar) vs Alcohol') +
       xlab("Alcohol") +
       ylab("Log10(Residual Sugar)")

p19 <- ggplot(aes(x = alcohol, y = log10(chlorides)), data = wines) + 
       geom_point(alpha = 1/5) +       
       geom_smooth(method = 'lm') +
       ggtitle('Log10(Chlorides) vs Alcohol') +
       xlab("Alcohol") +
       ylab("Log10(Chlorides)")

p20 <- ggplot(aes(x = alcohol, y = total.sulfur.dioxide), data = wines) +
       ylim(c(0, 300)) +
       geom_point(alpha = 1/5) +       
       geom_smooth(method = 'lm') +
       ggtitle('Total.sulfur.dioxide vs Alcohol') +
       xlab("Alcohol") +
       ylab("Total.sulfur.dioxide")

p21 <- ggplot(aes(x = alcohol, y = free.sulfur.dioxide), data = wines) +
       ylim(c(0, 90)) +
       geom_point(alpha = 1/5) +       
       geom_smooth(method = 'lm') +
       ggtitle('Free.sulfur.dioxide vs Alcohol') +
       xlab("Alcohol") +
       ylab("Free.sulfur.dioxide")

grid.arrange(p16,p17,p18,p19,p20,p21, ncol = 2)
```

We can see negative influence of these variables `density`, `residual.sugar`, `chlorides`, `total.sulfur.dioxide`, `free.sulfur.dioxide` in `alcohol` level. Only `pH` shows a positive correlation.

### pH vs fixed.acidity and pH vs residual sugar

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH, y = fixed.acidity), data = wines) + geom_jitter(alpha=1/5) + geom_smooth(method = 'lm') + ylim(c(3.8, 11)) 

cor(wines$pH, wines$fixed.acidity)

ggplot(aes(x = pH, y = residual.sugar), data = wines) + geom_jitter(alpha=1/5) + geom_smooth(method = 'lm') + ylim(c(0, 20)) 

cor.test(wines$pH, wines$residual.sugar, method='pearson')
```

I certainly would expect `fixed.acidity` to be correlated to `pH`. In chemistry when `acidity` goes down, the `pH` value should goes up as well. There's no clear pattern on `residual.sugar` and `pH`, even though they are negative correlated.

### free.sulfur.dioxide vs total.sulfur.dioxide

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), 
           data = subset(wines, free.sulfur.dioxide<quantile(wines$free.sulfur.dioxide,.99) & 
                  total.sulfur.dioxide<quantile(wines$total.sulfur.dioxide,.99))) +
  geom_point(alpha = 1/5) +
  geom_smooth(method='lm')

cor.test(wines$free.sulfur.dioxide, wines$total.sulfur.dioxide, method='pearson')
```

Besides main feature and secondary feature. Other features such as `free.sulfur.dioxide` and `total.sulfur.dioxide` are correlated to one another. 

### Residual Sugar vs Density
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = residual.sugar), data = 
           subset(wines, residual.sugar<quantile(wines$residual.sugar,.99) & density<quantile(density,.99))) +
        geom_point(alpha = 1/5) +       
        geom_smooth(method = 'lm') + 
        ggtitle('Residual Sugar vs Density') +
        xlab("Density") +
        ylab("Residual Sugar")
 
cor.test(wines$density, wines$residual.sugar, method='pearson')
```

When top 1% of observations is excluded and plotting both of these two variables `residual.sugar` and `density` together shows a strong correlation.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

Since our main goal was to find out what features affects `quality` score. The main relationship observed was `alcohol` and `quality`, `pH` and `quality`. Besides that `density` and `chlorides` have negative influence on `qualty`. 

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

The other features which shows interesting relationship was free.sulfur.dioxide and total.sulfur.dioxide with a correlation of 0.616. Both are correlated to one another as both are free and bound forms of sulfur dioxide gas (S02).

### What was the strongest relationship you found?

The strongest relationship found was density and residual.sugar with a correlation of 0.828.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(aes(x = density, y = residual.sugar), 
       data = subset(wines, residual.sugar<quantile(wines$residual.sugar,.99) & 
                  density<quantile(density,.99))) +
  geom_point(aes(color = alcohol)) + 
  geom_vline(xintercept = mean(wines$density), linetype='longdash', color='blue', alpha=.5) +
  geom_hline(yintercept = mean(wines$residual.sugar), linetype='longdash', color='blue', alpha=.5) +
  ggtitle('Residual Sugar vs Density') +
  xlab("Density") +
  ylab("Residual Sugar")

```

Since we found out that `density` and `residual.sugar` has the strongest relationship. I explored more with `alcohol`. It appears that as `density` and `residual.sugar` increases the `alcohol` color seems to turn darker in which less `alcohol`.


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = residual.sugar), 
       data = subset(wines, residual.sugar<quantile(wines$residual.sugar,.99) & 
                  density<quantile(density,.99))) +
        geom_jitter(alpha=.5, size=3, aes(size=quality_cut, color=quality_cut)) +
        geom_vline(xintercept = mean(wines$density), linetype='longdash', color='blue', alpha=.5) +
        geom_hline(yintercept = mean(wines$residual.sugar), linetype='longdash', color='blue', alpha=.5) +
        scale_color_brewer(type = 'div',
                       guide = guide_legend(title = 'Quality Cut', reverse = T)) +
        ggtitle('Residual Sugar vs Density') +
        xlab("Density") +
        ylab("Residual Sugar")
```

When we swap out alcohol with our defined variable `quality` cut. The result is more obvious. We can see that better quality score wines are concentrated on the left hand side of the plot where as bad quality score wines are on the right.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH, y = fixed.acidity), data = wines) + 
  geom_jitter(alpha=.3, aes(color=quality_as_factor)) + 
  guides(color=guide_legend(title="Quality"))

ggplot(aes(x = pH, y = fixed.acidity), data = wines) + 
  geom_jitter(alpha=.3, aes(color=quality_cut)) + 
  guides(color=guide_legend(title="Quality"))

```

Initially, I thought `pH` has a little bit of correlation with `quality`. Thus `pH` vs `fixed.acidity` should somewhat linked with `quality`. After plotting it out there seems no strong pattern that we can identify from `pH` vs `fixed.acidity` vs `quality`.

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?



### Were there any interesting or surprising interactions between features?

As I mentioned above, I thought `pH` would provide some interaction with `quality`. But after plotting it out. It's a bit hard to identify it from the plot.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x=alcohol), data = wines) +
  geom_bar(binwidth = 0.1) +
  xlab("Alcohol (% by volume)") +
  ylab("Occurence") +
  scale_x_continuous(limits = c(8,15), breaks = seq(8,15,1)) +
  ggtitle("Histogram for Alcohol Percentage")
```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data=wines, aes(x=alcohol, y=density)) +    
    geom_jitter(alpha=1/2, aes(size=quality_cut, color=quality_cut)) +
    scale_x_continuous(limits = c(quantile(wines$alcohol, 0.01), 
                                quantile(wines$alcohol, 0.99))) +
    scale_y_continuous(limits = c(quantile(wines$density, 0.01),
                                quantile(wines$density,0.99)))

ggplot(data=wines, aes(x=alcohol, y=pH)) + 
    geom_jitter(alpha=.5) +
    facet_wrap(~quality_as_factor) +
    geom_vline(xintercept = mean(wines$alcohol), linetype='longdash', color='blue', alpha=.5) +
    geom_hline(yintercept = mean(wines$pH), linetype='longdash', color='blue', alpha=.5)

ggplot(data=wines, aes(x=alcohol, y=pH)) +    
    geom_jitter(alpha=1/2, aes(size=quality_cut, color=quality_cut)) +  
    coord_cartesian(
        xlim=c(quantile(wines$alcohol,0),quantile(wines$alcohol,.99)),
        ylim=c(quantile(wines$pH,0),quantile(wines$pH,.99))
        )


ggplot(aes(x = density, y = residual.sugar), 
       data = subset(wines, residual.sugar<quantile(wines$residual.sugar,.99) & 
                  density<quantile(density,.99))) +
        geom_jitter(alpha=.5, size=3, aes(size=quality_cut, color=quality_cut))

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection


ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), 
           data = subset(wines, free.sulfur.dioxide<quantile(wines$free.sulfur.dioxide,.99) & 
                  total.sulfur.dioxide<quantile(wines$total.sulfur.dioxide,.99))) +
  geom_jitter(alpha=1/5, aes(color=density))
  

wines$quality.cut <- cut(wines$quality, breaks=c(0,3,5,7,10))
ggplot(data=wines, aes(x=alcohol, y=pH)) +
    coord_cartesian(
        xlim=c(quantile(wines$alcohol,.01),quantile(wines$alcohol,.99)),
        ylim=c(quantile(wines$pH,.01),quantile(wines$pH,.99))
        ) +
    geom_jitter(alpha=1/3, aes(size=quality.cut, color=quality_cut)) +
    geom_vline(xintercept = mean(wines$alcohol), linetype='longdash', color='blue', alpha=.5) +
    geom_hline(yintercept = mean(wines$pH), linetype='longdash', color='blue', alpha=.5)

ggplot(data=wines, aes(x=alcohol, y=pH, color=quality_as_factor)) +
    coord_cartesian(
        xlim=c(quantile(wines$alcohol,.01),quantile(wines$alcohol,.99)),
        ylim=c(quantile(wines$pH,.01),quantile(wines$pH,.99))
        ) +
    facet_wrap(~quality_cut) + 
    geom_point() +
    geom_vline(xintercept = mean(wines$alcohol), linetype='longdash', color='blue', alpha=.5) +
    geom_hline(yintercept = mean(wines$pH), linetype='longdash', color='blue', alpha=.5)
    
ggplot(aes(x = density, y = residual.sugar, color = quality_cut), data = 
    subset(wines, residual.sugar<quantile(wines$residual.sugar,.99) & density<quantile(density,.99))) +
    geom_point(alpha = 0.5, size = 2, position = 'jitter') +
    scale_color_brewer(type = 'div',
                       guide = guide_legend(title = 'Quality', reverse = T,
                                           override.aes = list(alpha = 1, size = 2))) +  
    scale_y_continuous(trans = log10_trans())

